---
date: "2020-12-07"
title: Carbon Footprint of Food Groups 
---

**DATASET 2: C02 EMISSION OF DIFFERENT FOOD TYPES**

For this project, I am looking at the carbon footprints of different categories of food. In this data set there are 11 different types of food looked at from 130 different countries. The rate of consumption by kg/person/year is also included in the data. The co2 emissions are measured in kg/person/year as well. 

**DATA**
```{r}
food_consumption <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv')
food_consumption
```

**KEY**


QUESTION 1: What is the average carbon dioxide emission for each food group? Which food groups were the highest?

I have always been told that animal products have the highest carbon footprint of anything that humans eat. The data below show this to be true. Beef had the highest carbon emission at almost 400 kg/person/year. Dairy products also had very high CO2 emissions at almost 200 kg/person/year.

AVERAGE CO2 EMISSION v.s. FOOD CATEGORY
```{r}
dt <- data.frame(readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv'))
library(ggplot2)
library(dplyr)
dt %>%
    group_by(food_category) %>%
    summarize(mean_co2 = mean(co2_emmission)) %>%
    ggplot(aes(x = food_category, y = mean_co2)) +
    geom_point() +
    ggtitle("Average Carbon Footprint of Food Groups") +
    coord_flip()
```

QUESTION 2: How much of each type of food is consumed per year? 

Dairy products were the most consumed food items. 

FOOD CATEGORY v.s. CONSUMPTION
```{r}
df <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv')
head(df)

library(ggplot2)

p<-ggplot(data=df, aes(x=food_category, y=consumption)) +
  geom_bar(stat="identity")
p
p + coord_flip()
```

QUESTION 3:How much co2 does each country create per year?

COUNTRY v.s. CO2
```{r}
dt <- data.frame(readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv'))
library(ggplot2)

dt %>%
    group_by(country) %>%
    summarize(total_co2 = sum(co2_emmission)) %>%
    ggplot(aes(x = country, y = total_co2)) +
    geom_point() +
    ggtitle("Carbon Dioxide Emission of Each Country") +
    coord_flip() 

```



QUESTION 4:How much co2 does the U.S. create in total each year? How much is created by each food group?

According to the data below, the United States creates 1718.86 kg/person/year of Carbon Dioxide through our food. Ethiopia has one of the lowest carbon footprints in this data, at about 241.92 kg/person/year, and that is because the country consumes many more plants than they do animals. 

TOTAL U.S C02 EMISSION 
```{r}
dt <- data.frame(readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-18/food_consumption.csv'))


dt %>%
    group_by(country) %>%
    summarize(total_co2 = sum(co2_emmission)) 
    
```



**CONCLUSION**

