---
date: "2020-12-07"
title: Passwords
---

**Dataset 1: Password Strength**

For this project, I am looking at a dataset that analyzes different passwords and their strengths. I got this data from TidyTuesday on Github. The passwords are grouped into 10 different categories. The data looks at how long each password took to crack and the unit of time associated with it. The data also looks at how long each password took to crack offline in seconds. The authors of the data also included the rank of the popularity of each password in their database. I thought this was really interesting to see what kinds of passwords other people are using. 

**DATA**
```{r}
passwords <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-14/passwords.csv')
passwords
```

**KEY**


QUESTION 1: Which category had the highest average strength?


CATEGORY v.s. AVERAGE STRENGTH

First, I wanted to look at how the category of password compared to the password strength. It is important for me to note that the password "strength" is relative and based off of the author's comparison and analysis of other "bad" passwords. It is a 1-10 scale with 10 being the highest and 1 being the lowest. Nonetheless, I thought it would be interesting to see how each category did in terms of password strength.
```{r}
df <- data.frame(readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-14/passwords.csv'))
library(ggplot2)
library(dplyr)
df %>%
    group_by(category) %>%
    summarize(mean_strength = mean(strength)) %>%
    ggplot(aes(x = category, y = mean_strength)) +
    geom_point() +
    ggtitle("Average Strengths of Each Passsword Category") +
    coord_flip()


```

QUESTION 2: How does the strength correlate with how long it took to crack offline?


PASSWORD STRENGTH v.s. HOW LONG IT TOOK TO CRACK OFFLINE IN SECONDS

I wanted to look at how the strength compared to how long it took them to crack the password offline. I kind of figured that the two would correlate, but I thought it might be interesting to see how it looked in graph form. 
```{r}
df <- data.frame(readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-14/passwords.csv'))
head(df)

library(ggplot2)

 ggplot(data=df, aes(x=strength, y=offline_crack_sec, group=1)) +
   geom_point() +
   ggtitle("How Long Did it Take to Crack Passwords Based on Rank")
```

QUESTION 3: Which category had the highest average rank?


CATEGORY v.s. AVERAGE RANK
```{r}
df <- data.frame(readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-01-14/passwords.csv'))
library(ggplot2)

df %>%
    group_by(category) %>%
    summarize(mean_rank = mean(rank)) %>%
    ggplot(aes(x = category, y = mean_rank)) +
    geom_point() +
    ggtitle("The Average Rank of Passwords") +
    coord_flip()
    
```



**CONCLUSION**
